---
title: "Hematological Clock 014" 
subtitle: "(Sex :`r params$sex`, Strain:`r params$strain`, Dataset:`r params$dataset`, Method to calculate Acceleration:`r params$method`) "

author: "Jorge Martinez-Romero"
date: "2024-01-5"
format: html 
params:
  sex: "All" # "F", "M", "All". 
  strain: "All" # "B6", "HET3", "DO", "All"
  method: "differences" # Calculate age acceleration as the "differences" between predicted age and chronological age or as "residuals" after regresing predicted age on chronological age using a non linear regresssion
  dataset: "All" #Select the dataset in where to perform analyses: "All","vald","validate2", "All_held_out","train"
toc: true   
self-contained: true  
---


## CLOCKS PERFORMANCE BY AGE RANGE: 
Stratify clock predictions by age decile

***
```{r setup, include=FALSE}
Sys.setlocale(category = "LC_ALL", locale = "Greek")
knitr::opts_chunk$set(echo = TRUE)
patho<-"/spin1/home/linux/martinezromerj2/0016_Hematological_clock_2023_Git/"
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, messages=FALSE, comment = NA, root.dir = patho, fig.pos = 'H')
options(max.print=2000,future.globals.maxSize = 4000 * 1024^5)
```

## Parameters
### Explain parameters.
- sex: "F".  Select sex: "F", "M", "All".
- strain: "All".  Select strain: "B6", "HET3", "DO", "All".
- method: "differences".  Calculate age acceleration as:.
  - "differences" between predicted age and chronological age or as
  - "residuals" after regressing predicted age on chronological age
    using a non linear loess.
- dataset: "All". Select the dataset in where to perform analyses:
    "All", "Vald" (Cohort C07 G08) "Validate2" (Pool of cohorts), "All_held_out" (Includes all samples held out during the training process [Vald + Validate2]), "train" (Training set only).
    
```{r libraries, echo=TRUE, include=FALSE}
extrafont::loadfonts(device="win")
library("ggplot2")
library("tidyverse")
library("dplyr")
library("survival") 
library("Metrics")
library("h2o")
library("data.table")
library("Metrics")
library("eha")
library("kableExtra")
library("quantreg")
library("exact2x2")
library("utils")
library("LaplacesDemon")
library("lme4")
library("DT")
library("splines")

```

###  Retrieve data and predict

```{r Retrieve Advia}
load(paste0(patho,"Tables/Transitions/ALL_merged_for_modeling_72_10_18_.Rdata",sep=""))
```
1)  Function add deltas

```{r}
Add_delta_variable<-function(df,V1,id){
  ftr<-data.table(df)
  V1<-V1
  V2 <- paste("Delta", V1, sep=".")
  BB<-ftr[, (V2) := .SD-shift(.SD, type='lag'), by = ID, .SDcols=V1]
  BB<-as.data.frame(BB)
  BB[is.na(BB)]<-0
  return(BB)
}
```

### Select features to add delta (No categoricals)

```{r}
Features<-c("rbc","mcv","mchc","mch","chcm","hdw"  
,"wbc","plt","glucose","bw")
```

------------------------------------------------------------------------

```{r}
Features
```

------------------------------------------------------------------------

### Load best DNN after tuning

```{r, echo=TRUE}
library(h2o)
h2o.init()
pat<-paste0(patho,"Models/Models_tunning/DNN_blood_62_N1691016783.zip",sep="")
imported_model <- h2o.import_mojo(pat)

```

### Predict using the best DNN tuned

```{r}
train$dataset<-"train"
validate2$dataset<-"validate2"
vald$dataset<-"vald"
data_all<-rbind(train,validate2,vald)
```


```{r}
res<-data_all %>% filter(dataset!="train")
res$age_month<-round(res$age_wk/1, digits = 0)
res$Decile <- cut(res$age_month, breaks = quantile(res$age_month, probs = seq(0, 1, 0.1)),
                   labels = FALSE)
res$Decile<-ntile(res$age_month,10)
res$Decile<-as.factor(res$Decile)
table(res$Decile)
```


```{r}
res<-Add_delta_variable(res,Features)
predict_DNN<-function(df){dt<-df
  Validate <- as.h2o(dt)
  Hematology_Based_Age<-h2o.predict(object = imported_model, newdata =Validate)
  Hematology_Based_Age<-as.vector(round(Hematology_Based_Age,digit=0))
  r_test<-cor.test(Hematology_Based_Age,dt$age_wk)
   samples <- dim(dt)[1]
  mice <- length(unique(dt$ID)) 
  max_age<-round(max(dt$age_wk), digits = 0)
  R<-round(r_test$estimate,digit=2)
  P<-round(r_test$p.value,digit=2)
  MAE<-round(mae(dt$age_wk,Hematology_Based_Age),digit=2)
  RMSE<-round(rmse(dt$age_wk,Hematology_Based_Age),digit=2)
  return(data.frame(R, P, MAE=MAE,RMSE=RMSE, Samples=samples, Mice=mice, To_age=max_age))
}
results_per_ageGroup<-lapply(split(res,res$Decile),predict_DNN)

Results<-do.call(rbind,results_per_ageGroup)
```

```{r,eval=FALSE}
write.csv(Results,paste0(patho,"Results_Cross.csv",sep=""))
RR<-res5 %>% select(2,4,29:666) %>% filter(dataset!="train") %>% as.data.frame()
colnames(RR)
```

```{r}
# Function to calculate MAE and RMSE
calculate_errors <- function(true_values, predicted_values) {
  mae_value <- mae(true_values, predicted_values)
  rmse_value <- rmse(true_values, predicted_values)
  return(data.frame(MAE = mae_value, RMSE = rmse_value))
}
```


```{r}
# Init Correlation, MAE, and RMSE tests
cor_results <- cor.test(RR$age_wk, RR$Age_out_rbc)
mae_rmse_results <- calculate_errors(RR$age_wk, RR$Age_out_rbc)
results_censor <- data.frame(
  Variable = "Init",
  Correlation = "Init",
  Pval = "Init",
  mae_rmse_results
)
```
